---
output: html_document
---
``` {r Knitr prep, cache=FALSE,include=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, dev='cairo_pdf', warning=FALSE)
```

# Charts based on data by grouping

```{r Load and prep}
setwd("~/github/local/PubFinCZ/")
source('./src/lib/lib_PubFinCZ_light.R')
library(pbtools)

# You can prep the data afresh from raw, specifying whether to exclude
# the two departments with discontinuities

# excludeMVaMZV <- FALSE
# source('./src/staff-super-data/SuperData_CompletePrepGrouping.Rmd')

# Or load the prepped data from file, with or without the exclusion
# using the DataFinalPrep() function

yearbreaks42003_2012 <- as.Date(c("2003-01-01", "2006-01-01",
                                  "2009-01-01", "2012-01-01"))
```

# Plots 

# Change in staff numbers by grp over time
``` {r Change in staff numbers by grp over time}

hh <- DataFinalPrep(exclusion = T)

title='Změna počtu zaměstnanců pod regulací rozpočtu 2003-2012'
ylab='Změna počtu zaměstnanců'
xlab=''
hh2 <- hh[(hh$variable=='Zam_skutecnost') &
            (hh$mingrpset) & !is.na(hh$value) &
            hh$value!=0,]
plot <- ggplot(hh2,aes(x=Year, y=perc_base-1, linetype=grp)) +
  geom_line(size=1.5) +
  scale_y_continuous(labels=percent) +
  labs(title=NULL, x=NULL, y=NULL) +
  guides(linetype=guide_legend(nrow=2,byrow=T)) +
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_linetype_manual(values=c(2,4,3,1)) +
  theme(legend.key.width=unit(2,'cm'))
plot
SavePlot(plotname = 'StaffChange_2003-12_grp_subsetBW',ploth=10, plotw=15,plotformat = 'png')
```

# Change in nominal pay bill by grp over time
``` {r Change in paybill by grp over time}

hh <- DataFinalPrep(exclusion = T)

title='Změna výdajů na platy 2003-2012'
ylab='2003 = 100% (změna výdajů na platy)'
xlab=''

hh2 <- hh[(hh$variable=='Platy_skutecnost') &
            (hh$mingrp) & !is.na(hh$value) &
            hh$value!=0,]
plot <- ggplot(hh2,aes(x=Year, y=perc_base-1, linetype=grp)) +
  geom_line(size=1.5) +
  scale_y_continuous(labels=percent) +
  labs(title=NULL, x=NULL, y=NULL) +
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_linetype_manual(values=c(2,4,3,1)) +
  theme(legend.key.width=unit(2,'cm')) +
  guides(linetype=guide_legend(nrow=2,byrow=T))
plot

SavePlot(plotname = 'PaybillChangeNom_2003-12_grp_subsetBW',ploth=10, plotw=15,
         plotformat = 'png')
```


# change in real paybill by grp over time
``` {r Change in real paybill by grp over time}

hh <- DataFinalPrep(exclusion = T)

hh2 <- hh[(hh$variable=='Platy_skutecnost') &
            (hh$mingrpset) & !is.na(hh$value) &
#            hh$exekutiva &
            hh$value!=0,]
plot <- ggplot(hh2,aes(x=Year, y=perc_base/Infl2003Base-1, linetype=grp)) +
  geom_line(size=1.5) +
  scale_y_continuous(labels=percent) +
  scale_linetype_manual(values=c(2,4,3,1)) +
  theme(legend.key.width=unit(2,'cm')) +
  guides(linetype=guide_legend(nrow=2,byrow=T)) +
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  labs(title=NULL, x=NULL, y=NULL)
SavePlot(plotname = 'PaybillChangeReal_2003-12_grp_subset',ploth=10, plotw=15,plotformat = 'png')
plot
```

# % real changes to avg salary by grp
``` {r % Real changes to avg salary by grp}

hh <- DataFinalPrep(exclusion = T)
hh2 <- hh[hh$mingrpset & hh$variable=='AvgSal_skutecnost',]

plot <- ggplot(hh2, aes(Year, realchange-1, colour=grpcolour, group=grp)) +
  geom_line(size=1.5) +
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_y_continuous(labels=percent) + 
  scale_colour_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='legend') +
  guides(colour=guide_legend(nrow=2,byrow=T)) + LoadCustomThemes(idcols)
plot

# SavePlot(plotname = 'SalaryChangeReal_2003-12_grp_subsetBW',ploth=10, plotw=15,plotformat = 'png')
```

# avg sal in 2013 prices
``` {r Avg salary in 2013 prices by group}

hh <- DataFinalPrep(exclusion = T)
hh2 <- hh[hh$mingrpset & hh$variable=='AvgSal_skutecnost',]

plot <- ggplot(hh2, aes(Year, value*Infl2013Base, colour=grpcolour, group=grp)) +
  geom_line(size=1.5) +
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_y_continuous(labels=space) +
  scale_colour_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='legend') +
  guides(colour=guide_legend(nrow=2,byrow=T)) +
  labs(title=NULL, x=NULL, y=NULL)
plot
SavePlot(plotname = 'Salary_2003-12_2013prices_grp_subset',ploth=10, plotw=15,plotformat = 'png')
```

# avg sal in nominal prices
``` {r Avg salary in nominal prices by group}

hh <- DataFinalPrep(exclusion = T)
hh2 <- hh[hh$maxgrpset & hh$variable=='AvgSal_skutecnost',]

plot <- ggplot(hh2, aes(Year, value, colour=grpcolour, group=grp)) +
  geom_line(size=1.5) +
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_y_continuous(labels=space) +
  scale_colour_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='legend') +
  guides(colour=guide_legend(nrow=2,byrow=T)) +
  labs(title=NULL, x=NULL, y=NULL)
plot
SavePlot(plotname = 'Salary_2003-12_nominal_grp_all',ploth=10, plotw=15,plotformat = 'png')
```


# paybill adjusted in nominal prices by grp, plan - AREA
if generating this graph, make sure nothing has been excluded (MV, MZV)
``` {r Paybill outturn in 2013 prices by grp}

hh <- DataFinalPrep(exclusion = F)
hh2 <- hh[hh$sgrp==FALSE & hh$Year!='2013-01-01' & hh$variable=='Platy_skutecnost',]
# hh2 <- hh[hh$sgrp==FALSE & hh$variable=='Platy_schvaleny',]
hh2 <- hh2[hh2$maxgrp,]

plot <- ggplot(hh2,aes(Year, value/1000000, fill=grpcolour, group=grp)) +
  geom_area(stat='identity',position='stack') +
#   scale_y_continuous(labels=c('0','5 mld', '10 mld', '15 mld', '20 mld'),
#                      breaks = c(0,5,10,15,20)) + 
  scale_y_continuous(labels=c('0','20 mld', '40 mld', '60 mld', '120 mld'),
                     breaks = c(0,20,40,60,120)) + 
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_fill_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='legend') +
  guides(fill=guide_legend(nrow=2,byrow=T)) +
  
  labs(title=NULL, x=NULL, y=NULL)
plot

SavePlot(plotname = 'Paybill_2003-12_nominal_grp_all_area',ploth=10, plotw=15,plotformat = 'png')
```

# paybill adjusted in 2012 prices by grp, plan - AREA
if generating this graph, make sure nothing has been excluded (MV, MZV)
``` {r Paybill outturn in 2013 prices by grp}

hh <- DataFinalPrep(exclusion = F)
hh2 <- hh[hh$sgrp==FALSE & hh$Year!='2013-01-01' & hh$variable=='Platy_skutecnost',]
# hh2 <- hh[hh$sgrp==FALSE & hh$variable=='Platy_schvaleny',]
hh2 <- hh2[hh2$mingrp,]

plot <- ggplot(hh2,aes(Year, value*Infl2012Base/1000000, fill=grpcolour, group=grp)) +
  geom_area(stat='identity',position='stack') +
  scale_y_continuous(labels=c('0','5 mld', '10 mld', '15 mld', '20 mld'),
                     breaks = c(0,5,10,15,20)) + 
#   scale_y_continuous(labels=c('0','20 mld', '40 mld', '60 mld', '120 mld'),
#                      breaks = c(0,20,40,60,120)) + 
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_fill_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='legend') +
  guides(fill=guide_legend(nrow=2,byrow=T)) +
  
  labs(title=NULL, x=NULL, y=NULL)
plot

SavePlot(plotname = 'Paybill_2003-13_2012prices_grp_subset_area',ploth=10, plotw=15,plotformat = 'png')
```

# paybill adjusted in 2012 prices by grp, plan - BAR
if generating this graph, make sure nothing has been excluded (MV, MZV)
``` {r Paybill outturn in 2013 prices by grp}
`
hh <- DataFinalPrep(exclusion = F)
hh2 <- hh[hh$sgrp==FALSE & hh$Year!='2013-01-01' & hh$variable=='Platy_skutecnost',]
# hh2 <- hh[hh$sgrp==FALSE & hh$variable=='Platy_schvaleny',]
hh2 <- hh2[hh2$mingrp,]

plot <- ggplot(hh2,aes(Year, value*Infl2012Base/1000000, fill=grpcolour, group=grp)) +
  geom_bar(stat='identity',position='stack') +
  scale_y_continuous(labels=c('0','5 mld', '10 mld', '15 mld', '20 mld'),
                     breaks = c(0,5,10,15,20)) + 
#   scale_y_continuous(labels=c('0','20 mld', '40 mld', '60 mld', '120 mld'),
#                      breaks = c(0,20,40,60,120)) + 
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_fill_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='legend') +
  guides(fill=guide_legend(nrow=2,byrow=T)) +
  labs(title=NULL, x=NULL, y=NULL)
plot

SavePlot(plotname = 'Paybill_2003-13_2012prices_grp_subset_bar',ploth=10, plotw=15,plotformat = 'png')
```

# paybill adjusted in 2012 prices by grp, plan - LINE
if generating this graph, make sure nothing has been excluded (MV, MZV)
``` {r Paybill outturn in 2013 prices by grp}

hh <- DataFinalPrep(exclusion = F)
hh2 <- hh[hh$sgrp==FALSE & hh$Year!='2013-01-01' & hh$variable=='Platy_skutecnost',]
# hh2 <- hh[hh$sgrp==FALSE & hh$variable=='Platy_schvaleny',]
hh2 <- hh2[hh2$mingrp,]

plot <- ggplot(hh2,aes(Year, value*Infl2012Base/1000000, colour=grpcolour, group=grp)) +
  geom_line(stat='identity', size=1.5) +
  scale_y_continuous(labels=c('0','5 mld', '10 mld', '15 mld', '20 mld'),
                     breaks = c(0,5,10,15,20)) + 
#   scale_y_continuous(labels=c('0','20 mld', '40 mld', '60 mld', '120 mld'),
#                      breaks = c(0,20,40,60,120)) + 
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_colour_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='legend') +
  guides(colour=guide_legend(nrow=2,byrow=T)) +
  labs(title=NULL, x=NULL, y=NULL)
plot

SavePlot(plotname = 'Paybill_2003-13_2012prices_grp_subset_line',ploth=10, plotw=15,plotformat = 'png')
```

# Staff gap by grp (%)
``` {r Staff % gap by grp}

hh <- DataFinalPrep(exclusion = T)
hh2 <- hh[hh$mingrpset & (hh$Year!='2013-01-01' & hh$variable=='Zam_upr2skut'),]
hh2 <- hh2[with(hh2, order(grp)), ]

plot <- ggplot(hh2, aes(Year, -(value-1), fill=grpcolour, group=grp)) +
  geom_bar(stat='identity',position='dodge') +
  scale_fill_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='none') +
  scale_y_continuous(labels=percent) + facet_wrap(~grp, nrow = 2) + 
  labs(title=NULL, x=NULL, y=NULL) +
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  labs(title=NULL, x=NULL, y=NULL)
plot

SavePlot(plotname = 'BudgetingStaffGap_percent_2003-12_grp_subset',ploth=10, plotw=15,plotformat = 'png')
```

# Paybill over/underspend by grp (%)
``` {r Paybill over/underspend % by grp}

hh <- DataFinalPrep(exclusion = T)
hh2 <- hh[hh$mingrp & (hh$Year!='2013-01-01' & hh$variable=='Platy_upr2skut'),]
hh2 <- hh2[with(hh2, order(grp)), ]

plot <- ggplot(hh2, aes(Year, -(value-1), fill=grpcolour, group=grp)) +

  geom_bar(stat='identity',position='dodge') +
  scale_fill_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='none') +
  scale_y_continuous(labels=percent) + facet_wrap(~grp,nrow = 2) + 
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  theme(axis.text.x=element_text(angle=90,vjust=.5,hjust=0)) +
  labs(title=NULL, x=NULL, y=NULL)
plot

SavePlot(plotname = 'BudgetingUnderspends_2003-12_grp_subset',ploth=10, plotw=15,plotformat = 'png')
```

# Staff gap by grp, abs
``` {r Staff gap by grp absolute}

hh <- DataFinalPrep(exclusion = T)
hh2 <- hh[hh$mingrpset & (hh$Year!='2013-01-01' & hh$variable=='Zam_uprMinusskut'),]
hh2 <- hh2[hh2$grp!='Vše' & hh2$grp!='Příspěvkové',]

plot <- ggplot(hh2, aes(Year, -value, fill=grpcolour, group=grp)) +
  geom_bar(stat='identity',position='dodge') +
  scale_fill_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='none') +
  scale_y_continuous(labels=space) + facet_wrap(~grp,nrow = 2) + 
  labs(title=NULL, x=NULL, y=NULL) +
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  theme(axis.text.x=element_text(angle=90,vjust=.5,hjust=0))
plot

SavePlot(plotname = 'BudgetingStaffGap_abs_grp_subset',ploth=10, plotw=15,plotformat = 'png')

```

# Salary 'raise' budgeting effect in CZK, by grp
``` {r Salary raise budgeting effect CZK}
hh <- DataFinalPrep(exclusion = T)
hh2 <- hh[hh$variable=='AvgSal_uprMinusskut' & hh$mingrpset,]
  
plot <- ggplot(hh2, aes(Year, -value, group=grp,fill=grpcolour)) +
  geom_bar(stat='identity') +
  scale_fill_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='none') +
  labs(title=NULL, x=NULL, y=NULL) +
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_y_continuous(labels=space) +
  facet_wrap(~grp,nrow=2) +
  theme(axis.text.x=element_text(angle=90,vjust=.5,hjust=0))
plot

SavePlot(plotname = 'BudgetingSalaryRaise_abs_grp_subset',ploth=10, plotw=15,plotformat = 'png')
```

# Salary as % of national/Prague, by grp
```{r Salary as % of national/Prague by grp}

hh <- DataFinalPrep(exclusion = T)
hh2 <- hh[(hh$variable=='AvgSal_skutecnost') & hh$mingrpset & !is.na(hh$value),]
hh2$relevavgsal <- hh2$czsal_all
hh2$relevavgsal[hh2$UO] <- hh2$phasal_all[hh2$UO]
hh2$grp2 <- paste0(hh2$grpcolour,hh2$variable)

plot <- ggplot(hh2,aes(x=Year, y=value/relevavgsal-1,
                       group=grp2)) +
  geom_hline(y=0,size=1,col='grey') +
  geom_line(size=1.5,aes(colour=grpcolour)) +
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_y_continuous(labels=percent) + 
  scale_colour_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='legend') +
  guides(colour=guide_legend(nrow=2,byrow=T)) +
  labs(title=NULL, x=NULL, y=NULL) + theme(axis.line.x=element_line())
plot

SavePlot(plotname = 'Sal_perc_of_national_grp_subset',ploth=10, plotw=15,plotformat = 'png')
```

# salary 'raise' budgeting effect in %, by grp
``` {r Salary raise budgeting effect %}
hh <- origdata
hh2 <- hh[hh$variable=='AvgSal_upr2skut' & hh$mingrpset,]

plot <- ggplot(hh2, aes(x=Year, y=-(value-1), fill=grpcolour)) +
  geom_bar(stat='identity') +
  scale_y_continuous(label=percent) +
  facet_wrap(~grp,nrow=2) + 
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_fill_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='none') +
  labs(title=NULL, x=NULL, y=NULL) +
  theme(axis.text.x=element_text(angle=90,vjust=.5,hjust=0))
plot

SavePlot(plotname = 'BudgetingSalaryRaise_percent_grp_subset',ploth=10, plotw=15,plotformat = 'png')

```

# Real salary changes, by group
``` {r Real salary change by grp}

hh <- DataFinalPrep(exclusion = T)
hh2 <- hh[hh$variable=='AvgSal_skutecnost' & hh$mingrpset,]

plot <- ggplot(hh2, aes(x=Year, y=realchange-1)) +
  geom_line(stat='identity',size=1.5) +
  scale_y_continuous(label=percent) +
  facet_wrap(~grp,nrow=2) + 
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_colour_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='none') +
  labs(title=NULL, x=NULL, y=NULL)
plot
```

# Nominal salary, by group
``` {r Real salary change by grp}

hh <- DataFinalPrep(exclusion = T)
hh2 <- hh[hh$variable=='AvgSal_skutecnost' & hh$mingrpset,]

plot <- ggplot(hh2, aes(x=Year, y=value,colour=grpcolour)) +
  geom_line(size=1.5) +
  scale_y_continuous(label=space) +
  scale_x_date(breaks=yearbreaks42003_2012,labels=date_format('%Y'),
               limits = c(as.Date("2003-01-01"), as.Date('2012-01-01'))) +
  scale_colour_identity(breaks=hh2$grpcolour,labels=hh2$grp,guide='legend') +
  guides(colour=guide_legend(breaks=hh2$grpcolour,labels=hh2$grp,nrow=2))
  labs(title=NULL, x=NULL, y=NULL)
plot
```